### 2022-04-25 manuscript/preambule.tex

%%%%%%%%%%TO FORMAT A PREPRINT%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[top=7cm,bottom=2.5cm,headheight=120pt,headsep=15pt,left=6cm,right=1.5cm,marginparwidth=4.2cm,marginparsep=0.5cm]{geometry}

\usepackage{marginnote}
\usepackage{ifthen}
\reversemarginpar  % sets margin notes to the left
\usepackage{lipsum} % Required to insert dummy text
\usepackage{calc}
\usepackage{siunitx}
\usepackage{lineno}
\usepackage{titlesec}
\usepackage{indentfirst}
%\usepackage[none]{hyphenat} % use only if there is a problem
% Use Unicode characters
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% Clean citations with biblatex
\usepackage[
backend=biber,
natbib=true,
sortcites=true,
defernumbers=true,
style=authoryear,
citestyle=authoryear-comp,
maxnames=99,
maxcitenames=2,
giveninits=true,
terseinits=true,
url=false
]{biblatex}
\DeclareNameAlias{default}{family-given}
\renewcommand*{\revsdnamepunct}{} % no comma between family and given names
\renewbibmacro{in:}{%
  \ifentrytype{article}{}{\printtext{\bibstring{in}\intitlepunct}}} % remove 'In:' before journal name
\DeclareFieldFormat[article]{pages}{#1} % remove pp.
\AtEveryBibitem{\ifentrytype{article}{\clearfield{number}}{}} % don't print issue numbers
\DeclareFieldFormat[article, inbook, incollection, inproceedings, misc, thesis, unpublished]{title}{#1} % title without quotes
\usepackage{csquotes}
\RequirePackage[english]{babel} % must be called after biblatex
\addbibresource{refs.bib}
\DeclareBibliographyCategory{ignore}
\addtocategory{ignore}{recommendation} % adding recommendation to 'ignore' category so that it does not appear in the References
% Clickable references. Use \url{www.example.com} or \href{www.example.com}{description} to add a clicky url
\usepackage{nameref}
\usepackage[pdfborder={0 0 0}]{hyperref}  % sets link border to white
\urlstyle{same}

% Include figures
\usepackage{graphbox}  % loads graphicx ppackage with extended options for vertical alignment of figures
% Line numbers
%\usepackage[right]{lineno}
% Improve typesetting in LaTex
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }
% Text layout and font (Open Sans)
\setlength{\parindent}{0.4cm}
\linespread{1.2}
\RequirePackage[default,scale=0.90]{opensans}
% Defining document colors
\usepackage{xcolor}
\definecolor{darkgray}{HTML}{808080}
\definecolor{mediumgray}{HTML}{6D6E70}
\definecolor{ligthgray}{HTML}{d9d9d9}
\definecolor{pciblue}{HTML}{74adca}
\definecolor{opengreen}{HTML}{77933c}
% Use adjustwidth environment to exceed text width
\usepackage{changepage}
% Adjust caption style
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,singlelinecheck=off]{caption}

% Headers and footers

% Headers and footers
\usepackage{fancyhdr}  % custom headers/footers
\pagestyle{fancy}  % enables customization of headers/footers
\fancyhfoffset[L]{4.5cm}  % offsets header and footer to the left to include margin
\renewcommand{\headrulewidth}{\ifnum\thepage=1 0.5pt \else 0pt \fi} % header ruler only on first page
%\renewcommand{\footrulewidth}{0.5pt}
\renewcommand{\footrulewidth}{\ifnum\thepage=1 0.5pt \else 0pt \fi}
% full logo on first page, then no logo on subsequent pages 
\lhead{\ifnum\thepage=1 \includegraphics[width=13.5cm]{\logoname}\else \fi}  
\chead{}
\rhead{}

\lfoot{\ifnum\thepage=1 \scriptsize \textsc{\color{mediumgray}\PCI} \else  \fi}
\cfoot{}
\rfoot{}
% End Headers and footers

% DOI's
\newcommand{\DOIlink}{\href{https://doi.org/\DOI}{https://doi.org/\DOI}}
\newcommand{\DOIrecommendationlink}{\href{https://doi.org/\DOIrecommendation}{https://doi.org/\DOIrecommendation}}

%which logo to display
\ifthenelse{\equal{\PCI}{Peer Community In Zoology}}{\newcommand{\logoname}{logo_PDF_zool.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Ecology}}{\newcommand{\logoname}{logo_PDF_ecology.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Evolutionary Biology}}{\newcommand{\logoname}{logo_PDF_evolbiol.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Animal Science}}{\newcommand{\logoname}{logo_PDF_animsci.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Archaeology}}{\newcommand{\logoname}{logo_PDF_archaeo.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Forest and Wood Sciences}}{\newcommand{\logoname}{logo_PDF_fws.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Genomics}}{\newcommand{\logoname}{logo_PDF_genomics.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Mathematical and Computational Biology}}{\newcommand{\logoname}{logo_PDF_mcb.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Network Science}}{\newcommand{\logoname}{logo_PDF_networksci.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Paleontology}}{\newcommand{\logoname}{logo_PDF_paleo.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Neuroscience}}{\newcommand{\logoname}{logo_PDF_neuro.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Ecotoxicology and Environmental Chemistry}}{\newcommand{\logoname}{logo_PDF_ecotoxenvchem.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Infections}}{\newcommand{\logoname}{logo_PDF_infections.jpg}}{}
\ifthenelse{\equal{\PCI}{Peer Community In Registered Reports}}{\newcommand{\logoname}{logo_PDF_rr.jpg}}{}

\newcommand{\beginingpreprint}{

\vspace*{0.5cm}
\begin{flushleft}
\baselineskip=30pt
% Margin information
\marginpar{
\large\textnormal{\color{pciblue}\\RESEARCH ARTICLE}\\
\vspace*{0.5pt}
\\
\includegraphics[align=c,width=0.5cm]{OAlogoblue.png} \space \large\textbf{\color{pciblue}Open Access}\\
\\
\includegraphics[align=c,width=0.5cm]{o_peerreview.png} \space \large\textbf{\color{pciblue}Open Peer-Review}\\
\\
\includegraphics[align=c,width=0.5cm]{o_data3.png} \space \large\textbf{\color{pciblue}Open Data}\\
\\
\includegraphics[align=c,width=0.5cm]{o_script3.png} \space \large\textbf{\color{pciblue}Open Code}\\
\\
\\
\\
\\
\raggedright
%\vspace*{3.25cm}
\scriptsize\textbf{Cite as:}\space
\citeas\\
\vspace*{0.5cm}
\textbf{Posted:} \datepub\\
\vspace*{0.5cm}
\textbf{Recommender:}\\
\recommender\\
\vspace*{0.5cm}
\textbf{Reviewers:}\\
\reviewers\\
\vspace*{0.5cm}
\textbf{Correspondence:}\\
\href{mailto:\email}{\email}\\

}
{\Huge
\fontseries{sb}\selectfont{\preprinttitle}}
\end{flushleft}
\vspace*{0.25cm}
\begin{flushleft}
%\marginpar{

%\marginnote{
%\raggedright
%\vspace*{3.25cm}
%\scriptsize\textbf{Cite as:}\space
%\fullcite{preprint}\\
%\vspace*{0.5cm}
%\textbf{Published:} \datepub\\
%\vspace*{0.5cm}
%\textbf{Recommender:}\\
%\recommender\\
%\vspace*{0.5cm}
%\textbf{Reviewers:}\\
%\reviewers\\
%\vspace*{0.5cm}
%\textbf{Correspondence:}\\
%\href{mailto:\email}{\email}\\
%\vspace*{0.5cm} 
%\vspace*{3cm}
%\vspace*{0.2cm}
%}[0.8cm]

\Large
\listauthors
\end{flushleft}
\bigskip
{\raggedright
\listinstitutions}
% Recommended preprint box
\begin{flushleft}
%\noindent
\fcolorbox{lightgray}{lightgray}{
\parbox{\textwidth - 2\fboxsep}{
\centering\large{\fontseries{sb}\selectfont{This article has been peer-reviewed and recommended by\\
\emph{\PCI} (\DOIrecommendationlink)}}\\
%\raggedright\large{\fontseries{sb}\selectfont{An article peer-reviewed and recommended by \emph{\PCI}, edited by \recommender \space on the basis of the reviews by \reviewers \space (DOI: \DOIrecommendationlink)}}\\
%\small \fullcite{preprint}
}}
\end{flushleft}
\vspace*{0.5cm}
\fcolorbox{pciblue}{pciblue}{
\parbox{\textwidth - 2\fboxsep}{
\vspace{0.25cm}
\textbf{\large{\textsc{Abstract}}}\\
\preprintabstract\\

\footnotesize{\textbf{\emph{Keywords: }}\preprintkeywords}
\vspace{0.25cm}}
}
\newpage
\newgeometry{margin=1in}
}