### 2022-05-11 slides/subsections/models.tex

\begin{frame}{Linear models - taxa}
    \begin{columns}[c] % The "c" option specifies centered vertical alignment while the "t" option is used for top vertical alignment

        \column{.6\textwidth} % Right column and width

        \begin{figure}
        \includegraphics[width=1\linewidth]{../figures/lms-split-2.png}
        \end{figure}

        \column{.4\textwidth} % Right column and width

        \textbf{Details}
        \begin{enumerate}
            \item Kids $> 18$ months old, species $> 10\%$ prevalence
            \item GLM: $bug \sim cogScore + age$
            \item There were ~15 bugs significant after FDR correction
        \end{enumerate}

    \end{columns}

\end{frame}