### 2022-06-01 manuscript/sections/results.tex

\subsection*{Outline}

\begin{itemize}
  \item Intial description: Cohort size, demographics, data types
  \item Basic analysis: Summary statistics, ordinations, permanovas, mantel tests
  \item Exploration of taxonomic profiles + cogscores / brain, cross-sectional. Linear models. \hl{Is this necessary??}
  \item Try HAllA on multivariate data?
  \item Predictive model (RF) - taxonomic profiles + cogscores.
  \item Taxonomic profiles + brain structure
  \item Gene functions + metabolomes
\end{itemize}