### 2023-02-17 manuscript/tex/main.tex
clean up

%%%%%%%%%%TO FORMAT A PREPRINT FOR A PCI%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper]{article}

\input{preamble.tex}

\title{Gut microbes and their genes or something else}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%  SET THE LIST OF AUTHORS WITH CORRESPONDING AFFILIATIONS  use \& before last author %%%%%%%%%%%%%%%%%%%%
\author{Kevin S. Bonham\textsuperscript{1}, \space
Guilherme Fahur Bottino\textsuperscript{1}, 
Firstname Lastname\textsuperscript{2}, \&
Vanja Klepac-Ceraj\textsuperscript{1}
%\fourthauthor \textsuperscript{i}
%etc...
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  SET THE ABSTRACT %%%%%%%%%%%%%%%%%

\begin{document}

\begin{abstract}
    The gastrointestinal tract, its resident microorganisms and the central
    nervous system are connected by biochemical signaling, also known as
    "microbiome-gut-brain-axis." Both the human brain and the gut microbiome
    have critical developmental windows in the first three years of life,
    raising the possibility that their development is co-occurring and
    likely co-dependent. Emerging evidence implicates gut microorganisms and
    microbiota composition in cognitive outcomes and neurodevelopmental
    disorders (e.g., autism and anxiety), but the influence of gut microbial
    metabolism on typical neurodevelopment has not been explored in detail.
    We investigated the relationship of the microbiome with the neuroanatomy
    and cognitive function of 361 healthy children, demonstrating that
    differences in gut microbial taxa and gene functions are associated with
    overall cognitive function as well as with differences in the size of
    multiple brain regions. Using a combination of multivariate linear
    models and machine learning (ML) models, we showed that many species,
    including \emph{Gordonibacter pamelae} and \emph{Blautia wexlerae}, were
    significantly associated with higher cognitive function, while some
    species such as \emph{Ruminococcus gnavus} were more commonly found in
    children with low cognitive scores regardless of age or maternal
    education. Microbial genes for enzymes involved in the metabolism of
    neuroactive compounds, particularly short-chain fatty acids such as
    glutamate and propionate, were also associated with cognitive function.
    In addition, ML models were able to use microbial taxa to predict the
    volume of brain regions, with particular taxa often dominating the
    feature importance metric for individual brain regions, such as \emph{B.
    wexlerae} being the most dominant for the parahippocampal region or
    GABA-producing \emph{Bacteroidetes ovatus} for left accumbens. These
    findings provide potential biomarkers of neurocognition and may lead to
    the future development of targets for early detection and early
    intervention.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%% Text of the preprint %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% use \emph{} for italics and \parencite{} to cite a reference and \ref{XX} to cite the \label{XX}
% use \\ and a blank line to mark the end of paragraph and starting of a new one

\section*{Introduction}

\input{sections/intro}

\section*{Results}

\input{sections/results}

\section*{Discussion}

\input{sections/discussion}

\section*{Material and methods}

\input{sections/methods}

\end{document}