### 2023-03-22 other/anpan.qmd
work on anpan

---
title: "Running Anpan"
format: html
---

## Installation

```{r}
#| eval: false
install.packages(c("ape", 
                   "data.table",
                   "dplyr", 
                   "fastglm",
                   "furrr", 
                   "ggdendro",
                   "ggnewscale",
                   "ggplot2",
                   "loo",
                   "patchwork",
                   "phylogram",
                   "posterior",
                   "progressr",
                   "purrr",
                   "R.utils",
                   "remotes",
                   "stringr",
                   "tibble",
                   "tidyselect"), Ncpus = 8) # add Ncpus = 4 to go faster

install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
```

```{r}
#| eval: false
library(cmdstanr)
check_cmdstan_toolchain()
install_cmdstan(cores = 4)
```

```{r}
#| eval: false
remotes::install_github("biobakery/anpan")
```

## Running

After identifying samples with *B. wexlerae* in our sample set,
I generated a table with stratified gene families by running

```fish
echo -e "sample\tfeature\tabundance" > data/bwexlerae_genefamilies.tsv
for s in (cat has_bwex.txt)
    fd "$s"_genefamilies /grace/sequencing/processed/mgx/humann/main/ -x rg s__Blautia_wexlerae | sd '^q' "$s\t" >> data/bwexlerae_genefamilies.tsv
end

sd '^.+\t(.+\t.+\t.+)$' '$1' data/bwexlerae_genefamilies.tsv
sd '^FG\d+_S\d+\t$' '' data/bwexlerae_genefamilies.tsv

```

and then added a header manually.

```{r}
library(data.table)

df <- fread("data/bwexlerae_genefamilies.tsv")
df2 <- dcast(df[, .(feature, sample, abundance)],
                 feature ~ sample,
                 value.var = "abundance",
                 fun.aggregate = sum)
names(df2)[1] = "# Gene Family"
```

```{r}
library(anpan)
future::plan(future::multicore, workers=16)
```

```{r}
bwex_results <- anpan(#bug_dir           = "/grace/sequencing/processed/mgx/humann/main",
                        bug_file          = "./data/bwexlerae_genefamilies_wide.tsv",
                        meta_file         = "./data/manuscript_meta.csv",
                        out_dir           = "./data/anpan",
                        # annotation_file   = "/path/to/annotation.tsv", #optional, used for plots
                        filtering_method  = "kmeans",
                        model_type        = "fastglm",
                        covariates        = c("ageMonths", "sex", "education"),
                        outcome           = "cogScore",
                        plot_ext          = "pdf",
                        save_filter_stats = TRUE)
```

```{r}
input_path = file.path("data", "anpan", "filter_stats",
                       "filtered_bwexlerae_genefamilies_wide.tsv")

model_input = fread(input_path)

plot_results(res         = bwex_results,
             model_input = model_input,
             covariates  = c("ageMonths", "sex", "education"), 
             outcome     = "cogScore",
             bug_name    = "Blautia wexlerae",
             cluster     = "both",
             show_trees  = TRUE)

```